%% =================================================================
%%  4.  The Verified Computational Engine
%% =================================================================
\section{The Verified Computational Engine}
\label{sec:computational_methodology}

%% -------------------------------------------------------------------
\subsection{Closing the Verification Gap via Lean FFI}
\label{subsec:lean_ffi}

A fundamental challenge in computational number theory is the
``verification gap'': translating abstract formal proofs into
executable, performant code without introducing unverified
implementation errors (e.g., floating-point inaccuracies or integer
overflows).  To eliminate this gap, the UALBF architecture relies on a
compiled Lean Foreign Function Interface (FFI) bridge
(\texttt{FFI.lean} and \texttt{lean\_\allowbreak{}ffi.rs}).

Instead of re-implementing validated mathematical functions in Rust, we
author executable definitions directly in Lean~4 (e.g.,
\texttt{ualbf\_\allowbreak{}compute\_\allowbreak{}sigma\_\allowbreak{}hi\_\allowbreak{}impl}
and
\texttt{ualbf\_\allowbreak{}mod\_\allowbreak{}inverse\_\allowbreak{}lo\_\allowbreak{}impl}).
These functions share the same trusted logical core as our proofs.
They are marked with \texttt{@[export]} and compiled into a static
C-core library.

The Rust orchestrator then binds to these formally verified routines
via FFI\@.  During execution, critical $\mathcal{O}(1)$ targeting
operations---such as the 128-bit modular inverse used in ray-casting
(\texttt{mod\_\allowbreak{}inverse\_\allowbreak{}128}) and the
evaluation of $\sigma(p^{2e})$ exact sums
(\texttt{compute\_\allowbreak{}sigma})---are transparently passed
through the FFI layer into the verified Lean native code.  This
seamless synergy between the Rust orchestrator and Lean~4 ensures that
the high-performance search engine always executes exactly the
arithmetic operations dictated and proven sound by the formal
specification, eliminating the possibility of a computationally derived
false positive.

This architecture integrates the FFI bridge as an executable boundary atop our
stratified four-layer Lean formalisation (\cref{sec:math_foundations}): the executable
definitions in \texttt{FFI.lean} depend on the Engine and QPN tiers for their
correctness guarantees, but are compiled into standalone C-callable
binaries that the Rust engine invokes without any Lean runtime overhead.


%% -------------------------------------------------------------------
\subsection{Phase~1: The Global Annihilation Sieve}
\label{subsec:global_sieve}

The UALBF computational engine is driven by a highly parallel, lock-free
Rust implementation that fuses the generation of candidate prefixes with
immediate algebraic validation.  The engine traverses an array of valid
prime power components $p^{2e}$ using a hybrid Depth-First Search (DFS)
architecture: at shallow tree depths (depth $\le 2$), work-stealing
concurrency via the Rayon framework spawns independent search spaces
across available CPU cores.  To eliminate memory allocation bottlenecks,
deeper traversal seamlessly switches to a sequential, zero-allocation
push/pop model.  Throughout the DFS, active branches are subjected to
aggressive mathematical sieves, dynamically tracking accumulated metrics
such as the running abundancy ratio
$\prod \sigma(p_i^{2e_i})/p_i^{2e_i}$ and enforcing strict divisibility
constraints to prune unviable subsets early.

Before any DFS traversal begins, the engine precomputes the universe of
valid prime power components via the \emph{Global Annihilation Sieve}
(\texttt{phase1\_\allowbreak{}global\_\allowbreak{}annihilation\_\allowbreak{}sieve}).
For each odd prime~$p$ up to 250{,}000 and exponent
$2e \in \{2, 4, 6, 8\}$, the engine computes $\sigma(p^{2e})$ using the
formally verified Lean FFI (\texttt{compute\_\allowbreak{}sigma}) and
factors it using cyclotomic polynomial decomposition.

\paragraph{Cyclotomic decomposition for efficient factorisation.}
Rather than factoring the full value $\sigma(p^{2e})$ directly---which
can exceed $10^{40}$---the engine exploits the cyclotomic expansion of
\cref{lem:cyclotomic-expansion}.  The routine
\texttt{factor\_\allowbreak{}sigma\_\allowbreak{}cyclotomic} computes
each cyclotomic factor $\Phi_d(p)$ for $d \mid (2e+1)$, $d > 1$, via
the M\"obius inversion identity
$\Phi_d(x) = \prod_{k \mid d} (x^k - 1)^{\mu(d/k)}$.  Each
$\Phi_d(p)$ is typically much smaller than the full $\sigma$ value,
enabling the Elliptic Curve Method (ECM) factoriser (via the
\texttt{prime\_\allowbreak{}factorization} crate) to complete
efficiently.  When a cyclotomic evaluation overflows $\mathtt{u128}$,
the engine falls back to factoring the full $\sigma$ value computed via
the Lean backend.

\paragraph{Mod-8 annihilation.}
Every prime factor~$q$ of $\sigma(p^{2e})$ is checked against the
modulo-8 obstruction (\cref{thm:mod8}): if $q \equiv 5$ or
$7 \pmod 8$, the component $p^{2e}$ is discarded.  By
\cref{thm:sieve-sound}, this pruning is sound---no legitimate QPN
factor is lost.

The surviving components are sorted by abundance ratio
$\sigma(p^{2e})/p^{2e}$ in descending order, placing the most
productive factors first for the DFS.

%% -------------------------------------------------------------------
\subsection{Depth-First Search for Prefix Construction}
\label{subsec:dfs_prefix}

The engine constructs candidate prefixes via a hybrid DFS over the
sieved components, mapping the architecture of
\texttt{dfs\_\allowbreak{}tree.rs} to the formalized exact valuation
constraints.

\paragraph{Hybrid parallelism.}
The combinatorial explosion of possible QPN prime factors necessitates a
dual-architecture traversal strategy:

\begin{itemize}
  \item \textbf{Shallow Depths (Rayon Work-Stealing):}
    At the root and initial layers of the search tree, the engine
    leverages the Rayon framework for concurrent, lock-free parallel
    execution.  Rayon's work-stealing scheduler saturates all available
    processor threads without thread contention.

  \item \textbf{Deep Traversals (Sequential Push/Pop):}
    Beyond depth~2, the active thread assumes exclusive execution,
    dynamically switching to a deterministic, zero-allocation sequential
    model: mutating a single pre-allocated stack of prime prefixes via
    simple \texttt{push} and \texttt{pop} primitives.  This keeps
    memory access constrained within the L1/L2 caches.
\end{itemize}

\paragraph{Running abundancy tracking.}
Each \texttt{Prefix} struct carries a
\texttt{current\_\allowbreak{}abundancy} field---the running product
$\prod_{i} \sigma(p_i^{2e_i})/p_i^{2e_i}$ over all components assigned
to the prefix.  This field is updated incrementally (multiply on push,
restore on pop) and drives several pruning mechanisms:

\begin{itemize}
  \item \textbf{Overflow Kill:}
    If $\texttt{current\_\allowbreak{}abundancy} > 2.000001$, the
    prefix is immediately discarded.  Since
    $H(N) = 2 + 1/N \approx 2$ for large~$N$, exceeding this
    threshold indicates an impossible abundancy overrun.

  \item \textbf{Suffix Starvation Kill:}
    Before each DFS extension, the engine checks whether
    $\texttt{current\_\allowbreak{}abundancy}
     \times \texttt{suffix\_\allowbreak{}abundance}[i]$
    can reach~2, where $\texttt{suffix\_\allowbreak{}abundance}[i]$ is
    a precomputed array storing the maximum achievable abundancy product
    from the remaining components.  If not, the branch is pruned
    (justified by \cref{thm:starvation}).

  \item \textbf{Yamada Tail Starvation:}
    Before ray-casting, the engine computes the theoretical maximum
    Euler product remaining for primes larger than the prefix's largest
    prime, using Yamada's sparse prime density limit for quasiperfect
    numbers~\cite{yamada2005}.  If
    $\texttt{current\_\allowbreak{}abundancy}
     \times \max\_\mathrm{tail} < 2$,
    the branch is killed.
\end{itemize}

%% -------------------------------------------------------------------
\subsection{Theoretical Frameworks for Deeper Horizons: Orchestration and Ray-Casting}
\label{subsec:ray_casting}

\emph{(Note: We transparently state that at the $10^{37}$ bound tested in this paper, shallow structural traps entirely exhausted the search space. Consequently, the deeper computational frameworks detailed in this section remained inactive during the current empirical run.)}

The ray-cast strategy serves as the core enumeration engine of the
UALBF\@.  Given a prefix~$N_L$, the algorithm must find valid suffix
candidates $N_R = z^2$.  From the AMBS target
(\cref{eq:ambs-target}), we have
$N_R \equiv -(2N_L)^{-1} \pmod{\sigma(N_L)}$.  Let $X_L$ be this
target residue.  The targeting step, executed via a fast 128-bit
modular inverse function
(\texttt{mod\_\allowbreak{}inverse\_\allowbreak{}128}) validated through
the Lean FFI, establishes that
$z^2 \equiv X_L \pmod{\sigma(N_L)}$.

To enumerate candidate roots~$z$, the engine employs the
Tonelli--Shanks quadratic root extraction algorithm, generalised for
composite moduli via the Chinese Remainder Theorem in the
\texttt{composite\_\allowbreak{}tonelli\_\allowbreak{}shanks} routine.
The algorithm factorises $\sigma(N_L)$ to extract a set of base roots
$\{r_i\}$.  Each root generates an arithmetic progression
$z = r_i + c\,\sigma(N_L)$ for $c \ge 0$, bounded by the global
search limits $z_{\min}$ and $z_{\max}$.

\paragraph{Sigma cache.}
To avoid recomputing $\sigma(p^{2e})$ inside the raycast inner loop,
the engine precomputes a hash-map cache (\texttt{SigmaCache}) keyed by
$(p, 2e)$ for all primes up to 250{,}000 and exponents up to~8.  Cache
misses fall back to the verified Lean backend.

\paragraph{Exact valuation sieving.}
Before any expensive factorisation is attempted on candidates~$z$, the
engine subjects them to rigorous exact sieving using precomputed
illegal valuations
(\texttt{generate\_\allowbreak{}illegal\_\allowbreak{}z\_\allowbreak{}valuations}).
This sieve precomputes prime power pairs $(p^e, p^{e+1})$ for which
$\sigma(p^{2e})$ violates the mod-8 obstruction
(\cref{thm:sieve-sound}).

For each candidate~$z$, the engine executes a $\mathcal{O}(1)$ exact
divisor state check: taking the remainder
$R \equiv z \pmod{p^{e+1}}$, if $R \equiv 0 \pmod{p^e}$ but
$R \neq 0$, then $\nu_p(z) = e$, implying $\nu_p(N_R) = 2e$ (since
$N_R = z^2$) and hence $p^{2e} \parallel N$.  By
\cref{lem:exact-val}, $\sigma(p^{2e})$ must divide $\sigma(N)$, but
computing $\sigma(p^{2e}) \pmod{8}$ yields an immediate parity
contradiction with \cref{thm:mod8}.  The candidate is proven
structurally impossible and discarded.


%% -------------------------------------------------------------------
\subsection{Theoretical Frameworks for Deeper Horizons: LLL Lattice Diophantine Pruning}
\label{subsec:lll_lattice_pruning}

\emph{(Note: We transparently state that at the $10^{37}$ bound tested in this paper, shallow structural traps entirely exhausted the search space. Consequently, the deeper computational frameworks detailed in this section remained inactive during the current empirical run.)}

To shape the top of the search tree early (specifically depths 3 to~6),
the engine employs Lenstra--Lenstra--Lov\'asz (LLL) lattice basis
reduction.  By analyzing the sum-of-divisors constraint as an additive
Diophantine equation over natural logarithms, the engine can
structurally prune branches without deep enumeration.

The abundancy contribution of each exact-valuation prime squared $p^2$
is bounded by its log-abundancy
$\ln(\sigma(p^2)/p^2) = \ln(1 + 1/p + 1/p^2)$.  To avoid
floating-point inaccuracies and ensure formal soundness, the engine
utilizes a logarithmic scaling trick.  Each logarithm is evaluated to
256-bit MPFR precision, scaled by an enormous factor of $2^{200}$, and
completely truncated to an exact $\mathbb{Z}$ value using the
\texttt{rug::Integer} crate.

For a candidate prefix composed of $k$ primes
$\{p_1, p_2, \dots, p_k\}$, the engine constructs a
$(k+1) \times (k+2)$ integer basis matrix~$B$.  The first $k$ rows
encode the scaled log-abundancies of the known prefix primes, while the
$(k+1)$-th row is a placeholder representing the minimum possible
contribution from the unknown suffix (modeled using the smallest prime
$p_{k+1} > p_k$).  Specifically, the matrix takes the form:
%
\[
  B =
  \begin{pmatrix}
    2^{200} & 0       & \cdots & 0       &
      \lfloor 2^{200} \ln(1 + \frac{1}{p_1} + \frac{1}{p_1^2}) \rfloor \\
    0       & 2^{200} & \cdots & 0       &
      \lfloor 2^{200} \ln(1 + \frac{1}{p_2} + \frac{1}{p_2^2}) \rfloor \\
    \vdots  & \vdots  & \ddots & \vdots  & \vdots \\
    0       & 0       & \cdots & 2^{200} &
      \lfloor 2^{200} \ln(1 + \frac{1}{p_{k+1}} + \frac{1}{p_{k+1}^2}) \rfloor
  \end{pmatrix}
\]

Applying exact LLL reduction over $\mathbb{Z}$ yields the shortest
basis vector, whose final component dictates the lattice's minimum
log-abundancy granularity---the smallest possible additive step these
specific log-abundancies can generate.

Simultaneously, the QPN equation $\sigma(N) = 2N + 1$ implies a
required target abundancy of exactly $2 + \frac{1}{N}$.  For any
combination of prime factors to satisfy this equation, their total
log-abundancy sum must fall within an incredibly narrow margin of
$\ln(2)$.  Mathematically, the tolerance bound is strictly bounded by
$\frac{1}{2N}$, leading to a scaled tolerance of
$\lfloor 2^{200} / (2N) \rfloor$.  If the absolute value of the
shortened vector's log-component strictly exceeds this scaled tolerance,
the lattice is insufficiently dense.  It becomes mathematically
impossible for any integer combination of these prime powers to close
the gap to $\ln(2)$, allowing the engine to safely and permanently
prune the prefix as structurally infeasible.

%% -------------------------------------------------------------------
\subsection{Lock-Free Concurrency \& Telemetry}
\label{subsec:concurrency_telemetry}

The computational pipeline utilizes the Rust Rayon crate for dynamic,
work-stealing concurrency.  To maximize parallel throughput and avoid
the thundering-herd problem, thread synchronization is managed strictly
via lock-free atomic states and read-write locks, minimizing blocking.
This lock-free architecture not only ensures CPU cores remain fully
saturated but also drives real-time telemetry, allowing external
monitoring tools to track the active primes and search progress
dynamically without introducing observational overhead.

%% -------------------------------------------------------------------
\subsection{Dynamic Minimum Factors}
\label{subsec:dynamic_minimum_factors}

Leveraging our formalized Prasad--Sunitha bound
(\cref{thm:prasad-sunitha}), the computational search floor dynamically
shifts during deep exploration.  When the depth-first search branch
definitively excludes 3 and~5 as prime factors of a candidate~$N$, the
engine's dynamic constraints automatically elevate the minimum required
distinct prime factors from the baseline of~7 up to~15.  This dynamic
adjustment drastically accelerates the search by instantly rejecting
branches that cannot possibly reach the required abundancy index.
